### 2015-05-26 paper.tex
Initial commit of reduced-length paper for DMAH

\documentclass{sig-alternate}

\usepackage[usenames,dvipsnames]{color}
\usepackage{xspace}

\usepackage{fancyvrb}
\DefineVerbatimEnvironment{code}{Verbatim}{fontsize=\small}
\DefineVerbatimEnvironment{example}{Verbatim}{fontsize=\small}

% Looks better (and more concise) than Times New Roman
\usepackage{times}
% Better spacing
\usepackage{microtype}
\usepackage[normalem]{ulem}
\usepackage{enumitem}
% Caption package both lets you set the spacing between figure and caption
% and also makes the \figref{} point to the right place.
\usepackage[font=bf,aboveskip=0pt,belowskip=-12pt]{caption}
% Get the caption package to work with ACM style
\DeclareCaptionType[placement=b,within=none]{copyrightbox}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newfont{\mycrnotice}{ptmr8t at 7pt}
\newfont{\myconfname}{ptmri8t at 7pt}
\let\crnotice\mycrnotice%
\let\confname\myconfname%

\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page. Copyrights for components of this work owned by others than the author(s) must be honored. Abstracting with credit is permitted. To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee. Request permissions from Permissions@acm.org.}
\conferenceinfo{}{DMAH 2015, September 4th, Kohala Coast, USA}
\copyrightetc{}
\crdata{}

% Allow comments by name
\newcommand{\note}[2]{
    \textbf{\textcolor{#1}{#2}}
}
\newcommand{\comment}[1]{\note{red}{#1}}
\newcommand{\dave}[1]{\note{PineGreen}{Dave: #1}}
\newcommand{\magda}[1]{\note{Cyan}{Magda: #1}}
\newcommand{\tom}[1]{\note{OrangeRed}{Tom: #1}}

% To remove all notes, uncomment the next line.
%\renewcommand{\note}[2]{\unskip}

\newcommand{\ie}{{\em i.e.}\ }
\newcommand{\eg}{{\em e.g.}\ }
\newcommand{\ea}{{\em et al.}\xspace}
\newcommand{\aka}{{\em a.k.a.}\ }

\usepackage{url}
\usepackage[pdftex]{hyperref}

% Figures
\newcommand{\figureimgs}{
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\linewidth]{figures/x_ray_image_montage}
  \vspace{-12pt}
  \caption{\label{fig:imgs}
  	X-ray diffraction images.
    A) An X-ray diffraction image exhibits $d_{10}$, $d_{11}$, and
    $d_{20}$ peaks; the first and last are proportional to the
    distance between adjacent rows of thick filaments. The $d_{10}$ is
    the primary peak being targeted for automated detection and
    fitting. B) In this image, two fibers at slight angles to each
    other have generated multiple $d_{10}$ peaks which must be
    segregated during the detection process.  
	}
	\vspace{-2pt}
\end{figure}
}

\newcommand{\figureworkflow}{
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\linewidth]{figures/img_analysis}
  \vspace{-8pt}
  \caption{\label{fig:workflow}
  	Analysis workflow.
    The blocked center region is detected and used for registration.
    Low signal regions and edges are masked. The image is smoothed and
    peaks are detected, clustered, and classified. The background
    (minus that surrounding diffraction lines) is computed and
    subtracted. Regions of interest surrounding the $d_{10}$ peaks are
    extracted and the generating distributions are fit with an MCMC
    sampler. We marginalize across peak parameters other than those of
    interest.  
	}
	\vspace{-2pt}
\end{figure}
}

\begin{document}
\title{Automated analysis of muscle X-ray diffraction \\ imaging with MCMC}
\author{C.\ David Williams$^{1,2}$,  Magdalena Balazinska$^2$, Thomas Daniel$^1$ \\
\affaddr{$^1$Department of Biology, $^2$Department of Computer Science \& Engineering}\\
\affaddr{University of Washington} \\
}
\maketitle

% \dave{Presumably this gets filled in later?}
%    \vspace{-8pt}
%    \category{H.2.4}{Database Management}{Systems---Distributed databases, Query processing, Relational databases}
%    \vspace{-8pt}
%    \keywords{Myria; parallel data management; Cloud service; astronomy}

\sloppy

%%%%%%%%%%%
\section{Questions addressed}
\label{sec:addressed}

In this work, we focus on the following questions:

\begin{itemize}[noitemsep]
\item Are sufficient conserved markers available in small-angle scattering diffraction images for automatic segmentation?
\item Does MCMC distribution fitting provide reliable estimates of diffraction peak properties such as center and decay?
\end{itemize}

%%%%%%%%%%%
\section{Motivation}
\label{sec:motivation}

Biologists have studied muscle structure and dynamics since antiquity
but it is only with the advent of X-ray diffraction techniques in the
1960s that it became possible to visualize the orientation and spacing
of the semicrystalline proteins that generate and transmit force
within muscle \cite{Millman1998}. In X-ray diffraction, structural
information is recorded when an X-ray beam passes through and scatters
off a muscle sample to produce an image such as that shown in
Figure~\ref{fig:imgs}.

Existing analytic approaches extract structural parameters using
expert hand-digitization with NIH ImageJ. This produces repeatable
measurements but remains subjective, fails to provide confidence
intervals for the measured values, and is not reproducible by naive
digitizers. Additionally hand digitization is a time-intensive
analytic technique, with a single digitizer able to process only a few
hundred images a day.  This has historically been sufficient, but new
high-speed imaging systems are allowing the investigation of
short-timescale components of muscle contraction and generating data
sets with many thousands of images. The need for an automated and
reproducible image analysis tool chain is clear.

We present the first components of a processing tool chain which first
segments the diffraction image into regions of interest using
conserved features and then samples the possible parameter values with
a Markov chain Monte Carlo approach.

%%%%%%%%%%%
\section{Prototype}
\label{sec:proto}

We initially focus on measuring the $d_{10}$ parameter, a crucial
spacing in muscle shown in Figure~\ref{fig:imgs}. The $d_{10}$ spacing
determines the distance which muscle's molecular motors must bridge in
order to bind and generate force \cite{Williams2013}. This distance
changes during contraction, regulating the force produced.

Images generated during experiments share several key features, which
serve as challenges or fiduciary marks during analysis. As seen in
Figure~\ref{fig:imgs}, the brightest part of the image background is
occluded by a circular stop. This physical block is put in place to
prevent damage to the detector from the high photon flux seen at the
center of the X-ray beam. Surrounding the blocked region, the
remainder of the image displays an exponentially decaying background.
Interrupting this largely symmetric background are the diffraction
peaks we locate and model.

\figureimgs

We apply our workflow to a test corpus of 1,220 images generated using
X-ray diffraction during insect flight muscle research at Argonne
National Laboratory's BioCAT Beamline. Sample high-quality and
challenging images are shown in Figures~\ref{fig:imgs}A and
\ref{fig:imgs}B. Our technique segments these images into regions of
interest, identifies the approximate locations of the diffraction
peaks, subtracts background from the regions surrounding the peaks,
and computes the distribution of possible parameters which underly the
shown peaks. This allows us to calculate peak-to-peak distances to
sub-pixel accuracy with a confidence interval of 90\%.

Because the images in our corpus are a standard sample of those
produced by high-speed X-ray diffraction, our positive preliminary
results are a strong indication for the potential of this approach.

\subsection{Image segmentation}

The central circular blocked region is first located and acts as a
relative reference point for subsequent operations. Consistent with
experimental design, it is assumed to contain the center of the
diffraction pattern, although the center of the diffraction pattern is
not assumed to be at the center of the blocked region. The central
blocked area is as dark or darker than the background at image edges
and the edge of the background surrounding it is one of the brightest
regions, so the image is first split into areas with values less than
and areas with values greater than two standard deviations above the
mean. This partitioning yields a binary image where the center blocked
region is surrounded by a halo that is the upper end of the pattern
background and occasional dots where diffraction peaks rise more than
two standard deviations above background. This binary image is
converted to a hierarchical contour set with OpenCV. The blocked
region is taken as the inner-most contour and then modeled as the
smallest circle which can contain it. This identification is
successful in greater than 99\% of images in our test corpus of 1,220
images.

With the central blocked region identified (shown as a dark magenta
overlay in Figure~\ref{fig:workflow}) we identify local maxima and
mask out those found in regions of the image unlikely to provide
diffraction peaks of interest.  After smoothing through convolution
with a Gaussian kernel maxima are identified by treating each 3 by 3
pixel region as a group to prevent stochastic brightness variation
from identifying maxima separated by single pixel gaps. The resulting
local maxima are then accepted as peaks or rejected based on masking
parameters.  We begin with a complete image and first mask a circular
zone centered on the central blocked region. This prevents the edge
between the blocked region and the background from being interpreted
as ring of peaks. Next we mask regions below the 80th percentile,
eliminating maxima in regions dominated by detector noise. Finally, we
mask regions near the image edge to cull partially cropped peaks. The
resulting unmasked area from which we keep maxima is depicted as light
purple overlay in Figure~\ref{fig:workflow}.

\figureworkflow

Next, maxima are clustered into pairs of peaks based on their distance
and angle from the center of the blocked region. Starting with those
maxima nearest the blocked region, a corresponding maxima is sought an
equal distance away from the blocked region and located such that the
angle formed by the two maxima and the center of the blocked region is
180$^\circ$. If no matched maxima exists within a 10\% tolerance, the
maxima is discarded as a spurious peak. With peak pairs now identified
(shown as color matched dots in Figure~\ref{fig:workflow}), the
diffraction center is identified by taking the mean location between
peak pairs and the background is fit and subtracted.

The background is subtracted by first masking arcs encompassing peak
pairs and then fitting a double exponential to a radial profile of the
remaining image. An arc swept 12$^\circ$ out on either side of each
peak pair is sufficient to mask the effect of the diffraction peaks on
the background to be generated (shown as a light gray arc under the
peak pairs in Figure~\ref{fig:workflow}). We next calculate a radial
profile of the image, centered on the pattern center and omitting the
regions surrounding diffraction lines. We fit a double exponential
function of the form $background = a+ b e^{-x c} + d e^{-x e}$ to the
radial profile. From these parameters we generate an estimates
background image and subtract it from the real image, allowing us to
extract the $d{10}$ peaks as regions of interest (ROIs) unhindered by
an overlaid diffraction background.

\subsection{Image modeling with MCMC processes}

With the background subtracted and the $d_{10}$ peaks identified and
isolated from the rest of the image as ROIs, we apply Markov chain
Monte Carlo (MCMC) sampling to determine the probability distributions
from which the peak parameters could be drawn. We treat the peaks as
being drawn from an underlying Pearson VII distribution, commonly used
to fit X-ray diffraction peaks. This process allows us to generate
possible peak matches using five parameters: peak center x-location,
peak center y-location, peak height, peak spread, and peak decay rate.
We perform an initial peak fitting by residual minimization between a
generated peak and the extracted ROI. This gives us a set of starting
parameters that we use, with random variation, to initialize the
positions of the MCMC agents that will explore the model space.

Before MCMC sampling we must define our query's likelihood and prior.
We choose a flat prior as our initial information about the model is
minimal. To calculate the likelihood we represent each pixel's photon
count as a Poisson process in the form $P(d|m)=e^{-m}
\left(m^{d}/d!\right)$ where $m$ is the model value and $d$ is the
experimental data value. These functions are fed into \textit{emcee},
an efficient MCMC analysis Python library \cite{ForemanMackey2013}.
After a burn in period of 100 steps, the sampler histories are erased
and a further 1000 steps are run to generate the posterior probability
distributions of our peak parameters.

One of MCMC modeling's convenient features is that extracting only a
subset of parameters marginalizes across those we discard. That is,
when we are interested in only the x- and y-locations of the peak
center to precisely calculate $d_{10}$ spacing (as in
Figure~\ref{fig:workflow}), we automatically integrate our uncertainty
about peak height, spread, and decay. After marginalizing over these
other parameters we obtain sub-pixel accuracy with a 95\% confidence
interval for the corpus images on which this technique has been
validated.  MCMC sampling combined with image segmentation allows us
to precisely, accurately, and automatically locate the $d_{10}$ peak
center and thus calculate the lattice spacing measured by a
diffraction image to within 0.03 nm.

%%%%%%%%%%%
\section{Challenging problems}
\label{sec:challenges}

Our approach produces an automated, reproducible, objective estimate of relevant image parameters but the following challenges remain:

\begin{itemize}[noitemsep]
\item Development of a declarative language to describe processing steps will speed use of this technique and ease reproducibility. 
\item Packaging of this tool chain into a cloud deployable
containerized image will enable trivial scaling to work with larger
datasets. 
\item Application of these techniques to coming ultra-high temporal-resolution images with far lower signal:noise will strain autosegmentation and peak-fitting techniques. 
\end{itemize}

%%%%%%%%%%%
%\section{Acknowledgments}
%\label{sec:acknowledgments}
%
%This work was supported in part by the Army Research Office through
%ARO Grants W911NF-13-1-0435 and W911NF-14-1-0396, an award from 
%the Gordon and Betty Moore Foundation and the Alfred P Sloan 
%Foundation, the Washington Research Foundation Fund for Innovation 
%in Data-Intensive Discovery, and the UW eScience Institute.
%
%We thank Jake VanderPlas for helpful discussions of statistical
%techniques, Tom Irving for advice on X-ray imaging, and Simon Sponberg
%for the sharing of diffraction images.

\scriptsize
\bibliography{articles,paper}
\bibliographystyle{abbrv}

\end{document}

### 2015-05-27 paper.tex
done with pass.

@@
 \begin{document}
-\title{Automated analysis of muscle X-ray diffraction \\ imaging with MCMC}
+\title{Automated Analysis of Muscle X-ray Diffraction \\ Imaging with MCMC}
 \author{C.\ David Williams$^{1,2}$,  Magdalena Balazinska$^2$, Thomas Daniel$^1$ \\
@@
 %%%%%%%%%%%
-\section{Questions addressed}
-\label{sec:addressed}
+%\section{Questions addressed}
+%\label{sec:addressed}
 
-In this work, we focus on the following questions:
+%In this work, we focus on the following questions:
 
-\begin{itemize}[noitemsep]
-\item Are sufficient conserved markers available in small-angle scattering diffraction images for automatic segmentation?
-\item Does MCMC distribution fitting provide reliable estimates of diffraction peak properties such as center and decay?
-\end{itemize}
+%\begin{itemize}[noitemsep]
+%\item Are sufficient conserved markers available in small-angle scattering diffraction images for automatic segmentation?
+%\item Does MCMC distribution fitting provide reliable estimates of diffraction peak properties such as center and decay?
+%\end{itemize}
 
@@
 
-Existing analytic approaches extract structural parameters using
-expert hand-digitization with NIH ImageJ. This produces repeatable
+Today, human experts manually extract structural parameters
+from X-ray images using NIH ImageJ. This produces repeatable
 measurements but remains subjective, fails to provide confidence
@@
 
-We present the first components of a processing tool chain which first
-segments the diffraction image into regions of interest using
+The goal of our project is to build a service for the automated
+analysis of X-ray images of muscle structures.  Users should specify
+the analysis that they need using a declarative query interface and
+the system should automatically process the user's image database.  In
+this extended abstract, we present the first components of the
+processing tool chain at the heart of this service. The toolchain
+first segments the diffraction image into regions of interest using
 conserved features and then samples the possible parameter values with
@@
 Interrupting this largely symmetric background are the diffraction
-peaks we locate and model.
+peaks we locate and model.  Our core data analysis pipeline includes
+two steps: image segmentation and image modeling with MCMC processes.
 
 \figureimgs
-
-We apply our workflow to a test corpus of 1,220 images generated using
-X-ray diffraction during insect flight muscle research at Argonne
-National Laboratory's BioCAT Beamline. Sample high-quality and
-challenging images are shown in Figures~\ref{fig:imgs}A and
-\ref{fig:imgs}B. Our technique segments these images into regions of
-interest, identifies the approximate locations of the diffraction
-peaks, subtracts background from the regions surrounding the peaks,
-and computes the distribution of possible parameters which underly the
-shown peaks. This allows us to calculate peak-to-peak distances to
-sub-pixel accuracy with a confidence interval of 90\%.
-
-Because the images in our corpus are a standard sample of those
-produced by high-speed X-ray diffraction, our positive preliminary
-results are a strong indication for the potential of this approach.
 
@@
 region is taken as the inner-most contour and then modeled as the
-smallest circle which can contain it. This identification is
-successful in greater than 99\% of images in our test corpus of 1,220
-images.
+smallest circle that can contain it.
 
+\magda{This paragraph is tedious and somewhat confusing but I'm
+not sure how best to fix it.}
 With the central blocked region identified (shown as a dark magenta
@@
 
+\magda{Before we can say the following, we should say that peaks
+come in pairs.}
 Next, maxima are clustered into pairs of peaks based on their distance
@@
 peak pairs in Figure~\ref{fig:workflow}). We next calculate a radial
-profile of the image, centered on the pattern center and omitting the
-regions surrounding diffraction lines. We fit a double exponential
-function of the form $background = a+ b e^{-x c} + d e^{-x e}$ to the
-radial profile. From these parameters we generate an estimates
-background image and subtract it from the real image, allowing us to
-extract the $d{10}$ peaks as regions of interest (ROIs) unhindered by
-an overlaid diffraction background.
+profile of the image, \magda{centered on the center sounds repetitive}
+centered on the pattern center and omitting the regions surrounding
+diffraction lines. We fit a double exponential function of the form
+$background = a+ b e^{-x c} + d e^{-x e}$ to the radial profile. From
+these parameters we generate an estimates background image and
+subtract it from the real image, allowing us to extract the $d{10}$
+peaks as regions of interest (ROIs) unhindered by an overlaid
+diffraction background.
 
@@
 Figure~\ref{fig:workflow}), we automatically integrate our uncertainty
-about peak height, spread, and decay. After marginalizing over these
-other parameters we obtain sub-pixel accuracy with a 95\% confidence
-interval for the corpus images on which this technique has been
-validated.  MCMC sampling combined with image segmentation allows us
-to precisely, accurately, and automatically locate the $d_{10}$ peak
-center and thus calculate the lattice spacing measured by a
+about peak height, spread, and decay.
+
+\section{Preliminary Evaluation}
+
+We apply our workflow to a test corpus of 1,220 images generated using
+X-ray diffraction during insect flight muscle research at Argonne
+National Laboratory's BioCAT Beamline. Sample high-quality and
+challenging images are shown in Figures~\ref{fig:imgs}A and
+\ref{fig:imgs}B. Our technique segments these images into regions of
+interest, identifies the approximate locations of the diffraction
+peaks, subtracts background from the regions surrounding the peaks,
+and computes the distribution of possible parameters which underly the
+shown peaks.  We find that our initial image segmentation step
+successfully models the center blocked region of each image in greater
+than 99\% of images in our test corpus of 1,220 images. The overall
+process allows us to calculate peak-to-peak distances to sub-pixel
+accuracy with a confidence interval of 90\%.  After marginalizing over
+these other parameters we obtain sub-pixel accuracy with a 95\%
+confidence interval for the corpus images on which this technique has
+been validated.  MCMC sampling combined with image segmentation allows
+us to precisely, accurately, and automatically locate the $d_{10}$
+peak center and thus calculate the lattice spacing measured by a
 diffraction image to within 0.03 nm.
 
+Because the images in our corpus are a standard sample of those
+produced by high-speed X-ray diffraction, our positive preliminary
+results are a strong indication for the potential of this approach.
+
 %%%%%%%%%%%
-\section{Challenging problems}
+\section{Challenging and Next Steps}
 \label{sec:challenges}
 
-Our approach produces an automated, reproducible, objective estimate of relevant image parameters but the following challenges remain:
+Our initial data processing pipeline produces an automated,
+reproducible, objective estimate of relevant image parameters but the
+following challenges remain:
 
@@
 \item Development of a declarative language to describe processing steps will speed use of this technique and ease reproducibility. 
+The key question is to define the types of operations that users should be able to specify and how to specify them. Our goal is
+to generalize to a broad set of analysis needs for X-ray images of muscle structures.
 \item Packaging of this tool chain into a cloud deployable
 containerized image will enable trivial scaling to work with larger
-datasets. 
+datasets. Furthermore, the ability to access to tool chain directly through a web browser 
+with automatic back-end deployment of the analysis pipeline will facilitate adotpion. 
 \item Application of these techniques to coming ultra-high temporal-resolution images with far lower signal:noise will strain autosegmentation and peak-fitting techniques. 

### 2015-05-27 paper.tex
Incorporation of Magda's edits and recommendations

@@
 remainder of the image displays an exponentially decaying background.
-Interrupting this largely symmetric background are the diffraction
-peaks we locate and model.  Our core data analysis pipeline includes
-two steps: image segmentation and image modeling with MCMC processes.
+Interrupting this background are symmetric pairs of diffraction peaks
+we locate and model.  Our core data analysis pipeline includes two
+steps: image segmentation and image modeling with MCMC processes.
 
@@
 
-The central circular blocked region is first located and acts as a
-relative reference point for subsequent operations. Consistent with
+The dark central circular blocked region is first located and acts as
+a relative reference point for subsequent operations. Consistent with
 experimental design, it is assumed to contain the center of the
 diffraction pattern, although the center of the diffraction pattern is
-not assumed to be at the center of the blocked region. The central
-blocked area is as dark or darker than the background at image edges
-and the edge of the background surrounding it is one of the brightest
-regions, so the image is first split into areas with values less than
-and areas with values greater than two standard deviations above the
-mean. This partitioning yields a binary image where the center blocked
-region is surrounded by a halo that is the upper end of the pattern
-background and occasional dots where diffraction peaks rise more than
-two standard deviations above background. This binary image is
-converted to a hierarchical contour set with OpenCV. The blocked
-region is taken as the inner-most contour and then modeled as the
-smallest circle that can contain it.
+not assumed to be at the center of the blocked region. The edge of the
+background surrounding the block is the brightest region, so the image
+is first split between areas with values less than and greater than
+two standard deviations above the mean. This partitioning yields a
+binary image where the center blocked region is surrounded by a halo
+of the upper end of the pattern background and occasional dots where
+diffraction peaks rise more than two standard deviations above
+background. This binary image is converted to a hierarchical contour
+set with OpenCV. The blocked region is taken as the inner-most contour
+and then modeled as the smallest enclosing circle, shown as dark
+magenta in Figure~\ref{fig:workflow}.
 
-\magda{This paragraph is tedious and somewhat confusing but I'm
-not sure how best to fix it.}
-With the central blocked region identified (shown as a dark magenta
-overlay in Figure~\ref{fig:workflow}) we identify local maxima and
-mask out those found in regions of the image unlikely to provide
-diffraction peaks of interest.  After smoothing through convolution
-with a Gaussian kernel maxima are identified by treating each 3 by 3
-pixel region as a group to prevent stochastic brightness variation
-from identifying maxima separated by single pixel gaps. The resulting
-local maxima are then accepted as peaks or rejected based on masking
-parameters.  We begin with a complete image and first mask a circular
-zone centered on the central blocked region. This prevents the edge
-between the blocked region and the background from being interpreted
-as ring of peaks. Next we mask regions below the 80th percentile,
-eliminating maxima in regions dominated by detector noise. Finally, we
-mask regions near the image edge to cull partially cropped peaks. The
+\dave{I cut this down and simplified. It can be removed or further
+reworked if desired.}
+With the central blocked region located, we identify local maxima in 3
+by 3 pixel groups after Gaussian smoothing and mask those found in
+regions unlikely to provide peaks of interest.  Resulting maxima are
+accepted or rejected as peaks based on masking parameters.  Masked
+rejection-regions consist of: 1) a circular zone around the central
+blocked region where the blocking generates non-peak local maxima, 2)
+areas below the 80th percentile where detector noise dominates, 3)
+areas near the image edge where peaks are partially cropped. The
 resulting unmasked area from which we keep maxima is depicted as light
@@
 
-\magda{Before we can say the following, we should say that peaks
-come in pairs.}
-Next, maxima are clustered into pairs of peaks based on their distance
+Next, maxima are clustered into peak pairs based on their distance
 and angle from the center of the blocked region. Starting with those
@@
 remaining image. An arc swept 12$^\circ$ out on either side of each
-peak pair is sufficient to mask the effect of the diffraction peaks on
-the background to be generated (shown as a light gray arc under the
-peak pairs in Figure~\ref{fig:workflow}). We next calculate a radial
-profile of the image, \magda{centered on the center sounds repetitive}
-centered on the pattern center and omitting the regions surrounding
-diffraction lines. We fit a double exponential function of the form
-$background = a+ b e^{-x c} + d e^{-x e}$ to the radial profile. From
-these parameters we generate an estimates background image and
-subtract it from the real image, allowing us to extract the $d{10}$
-peaks as regions of interest (ROIs) unhindered by an overlaid
-diffraction background.
+peak pair masks the effect of the diffraction peaks on the background
+(shown as a light gray arc under the peak pairs in
+Figure~\ref{fig:workflow}). We calculate a radial profile of the image
+around the pattern center, omitting diffraction line regions. We fit a
+double exponential function of the form $background = a+ b e^{-x c} +
+d e^{-x e}$ to the radial profile. From these parameters we generate
+an estimated background image and subtract it from the real image,
+allowing us to extract the $d{10}$ peaks as regions of interest (ROIs)
+unhindered by an overlaid diffraction background.
 
@@
 
+%%%%%%%%%%%
 \section{Preliminary Evaluation}
+\label{sec:eval}
 
@@
 process allows us to calculate peak-to-peak distances to sub-pixel
-accuracy with a confidence interval of 90\%.  After marginalizing over
-these other parameters we obtain sub-pixel accuracy with a 95\%
-confidence interval for the corpus images on which this technique has
-been validated.  MCMC sampling combined with image segmentation allows
-us to precisely, accurately, and automatically locate the $d_{10}$
-peak center and thus calculate the lattice spacing measured by a
-diffraction image to within 0.03 nm.
+accuracy with a confidence interval of 90\%. MCMC sampling combined
+with image segmentation allows us to precisely, accurately, and
+automatically locate the $d_{10}$ peak center and thus calculate the
+lattice spacing measured by a diffraction image to within 0.03 nm.
 
@@
 %%%%%%%%%%%
-\section{Challenging and Next Steps}
+\section{Challenges and Next Steps}
 \label{sec:challenges}
@@
 \begin{itemize}[noitemsep]
-\item Development of a declarative language to describe processing steps will speed use of this technique and ease reproducibility. 
-The key question is to define the types of operations that users should be able to specify and how to specify them. Our goal is
-to generalize to a broad set of analysis needs for X-ray images of muscle structures.
+\item Development of a declarative language to describe processing
+    steps will speed use of this technique and ease reproducibility.
+    The key question is to define the types of operations that users
+    should be able to specify and how to specify them. Our goal is to
+    generalize to a broad set of analysis needs for X-ray images of
+    muscle structure.
 \item Packaging of this tool chain into a cloud deployable
-containerized image will enable trivial scaling to work with larger
-datasets. Furthermore, the ability to access to tool chain directly through a web browser 
-with automatic back-end deployment of the analysis pipeline will facilitate adotpion. 
-\item Application of these techniques to coming ultra-high temporal-resolution images with far lower signal:noise will strain autosegmentation and peak-fitting techniques. 
+    containerized image will enable trivial scaling to work with
+    larger datasets. Furthermore, the ability to access to tool chain
+    directly through a web browser with automatic back-end deployment
+    of the analysis pipeline will facilitate adoption. 
+\item Application of these techniques to coming ultra-high
+    temporal-resolution images with far lower signal:noise will strain
+    autosegmentation and peak-fitting techniques. 
 \end{itemize}

### 2015-05-29 paper.tex
Incorporated Tom's edits, removed passive voice

@@
     $d_{20}$ peaks; the first and last are proportional to the
-    distance between adjacent rows of thick filaments. The $d_{10}$ is
-    the primary peak being targeted for automated detection and
-    fitting. B) In this image, two fibers at slight angles to each
-    other have generated multiple $d_{10}$ peaks which must be
-    segregated during the detection process.  
+    distance between adjacent rows of thick filaments. B) In this
+    image, two fibers at slight angles to each other have generated
+    multiple $d_{10}$ peaks which must be segregated during the
+    detection process.  
 	}
@@
   	Analysis workflow.
-    The blocked center region is detected and used for registration.
-    Low signal regions and edges are masked. The image is smoothed and
+    The blocked center region is detected and used as a landmark.  Low
+    signal regions and edges are masked. The image is smoothed and
     peaks are detected, clustered, and classified. The background
-    (minus that surrounding diffraction lines) is computed and
+    (sans that surrounding diffraction lines) is computed and
     subtracted. Regions of interest surrounding the $d_{10}$ peaks are
@@
 \title{Automated Analysis of Muscle X-ray Diffraction \\ Imaging with MCMC}
-\author{C.\ David Williams$^{1,2}$,  Magdalena Balazinska$^2$, Thomas Daniel$^1$ \\
+\author{C.\ David Williams$^{1}$,  Magdalena Balazinska$^2$, Thomas Daniel$^1$ \\
 \affaddr{$^1$Department of Biology, $^2$Department of Computer Science \& Engineering}\\
@@
 
-Biologists have studied muscle structure and dynamics since antiquity
-but it is only with the advent of X-ray diffraction techniques in the
-1960s that it became possible to visualize the orientation and spacing
-of the semicrystalline proteins that generate and transmit force
-within muscle \cite{Millman1998}. In X-ray diffraction, structural
-information is recorded when an X-ray beam passes through and scatters
-off a muscle sample to produce an image such as that shown in
-Figure~\ref{fig:imgs}.
+The molecular regulation of force generating protein systems remains a
+fundamental open problem in biology. Muscle uses a highly organized
+lattice of interacting elastic and force generating molecules to
+create controlled macro-scale movement \cite{Millman1998}. The advent
+of advanced X-ray diffraction imaging lets us analyze protein motions
+at spatial and temporal scales never previously realized, giving novel
+insight into the molecular basis of motion in living systems.
 
-Today, human experts manually extract structural parameters
-from X-ray images using NIH ImageJ. This produces repeatable
-measurements but remains subjective, fails to provide confidence
+Surprisingly, we still use human experts to manually extract
+structural parameters from X-ray images with NIH ImageJ. This produces
+repeatable measurements but is subjective, fails to provide confidence
 intervals for the measured values, and is not reproducible by naive
@@
 analytic technique, with a single digitizer able to process only a few
-hundred images a day.  This has historically been sufficient, but new
-high-speed imaging systems are allowing the investigation of
-short-timescale components of muscle contraction and generating data
-sets with many thousands of images. The need for an automated and
-reproducible image analysis tool chain is clear.
+hundred images a day.  Historically this rate has been sufficient, but
+new high-speed imaging systems let us investigate short-timescale
+components of muscle contraction and generate data sets with many
+thousands of images. The need for an automated and reproducible image
+analysis tool chain is clear.
 
-The goal of our project is to build a service for the automated
-analysis of X-ray images of muscle structures.  Users should specify
-the analysis that they need using a declarative query interface and
-the system should automatically process the user's image database.  In
-this extended abstract, we present the first components of the
-processing tool chain at the heart of this service. The toolchain
-first segments the diffraction image into regions of interest using
-conserved features and then samples the possible parameter values with
-a Markov chain Monte Carlo approach.
+We seek to build a service for the automated analysis of muscle
+structure X-ray images. Users should specify the analysis they need
+using a declarative query interface and the system should
+automatically process the user's image database. In this extended
+abstract, we present the first components of the processing tool chain
+at the heart of this service. The toolchain first segments the
+diffraction image into regions of interest using conserved features
+and then samples the possible parameter values with a Markov chain
+Monte Carlo approach.
 
@@
 
-We initially focus on measuring the $d_{10}$ parameter, a crucial
-spacing in muscle shown in Figure~\ref{fig:imgs}. The $d_{10}$ spacing
-determines the distance which muscle's molecular motors must bridge in
-order to bind and generate force \cite{Williams2013}. This distance
-changes during contraction, regulating the force produced.
+We initially focus on the $d_{10}$ parameter that determines the
+distance which muscle's molecular motors must bridge in order to bind
+and generate force \cite{Williams2013}. This distance changes during
+contraction and thus regulates the force produced.
 
@@
 Figure~\ref{fig:imgs}, the brightest part of the image background is
-occluded by a circular stop. This physical block is put in place to
-prevent damage to the detector from the high photon flux seen at the
-center of the X-ray beam. Surrounding the blocked region, the
-remainder of the image displays an exponentially decaying background.
-Interrupting this background are symmetric pairs of diffraction peaks
-we locate and model.  Our core data analysis pipeline includes two
+occluded by a circular stop. This physical block prevents damage to
+the detector from the high photon flux at the center of the X-ray
+beam. Surrounding the blocked region, the remainder of the image
+displays an exponentially decaying background. We must locate and
+model the symmetric pairs of diffraction peaks interrupting the
+exponential background. Our core data analysis pipeline includes two
 steps: image segmentation and image modeling with MCMC processes.
@@
 
-The dark central circular blocked region is first located and acts as
-a relative reference point for subsequent operations. Consistent with
-experimental design, it is assumed to contain the center of the
-diffraction pattern, although the center of the diffraction pattern is
-not assumed to be at the center of the blocked region. The edge of the
-background surrounding the block is the brightest region, so the image
-is first split between areas with values less than and greater than
-two standard deviations above the mean. This partitioning yields a
-binary image where the center blocked region is surrounded by a halo
-of the upper end of the pattern background and occasional dots where
-diffraction peaks rise more than two standard deviations above
-background. This binary image is converted to a hierarchical contour
-set with OpenCV. The blocked region is taken as the inner-most contour
-and then modeled as the smallest enclosing circle, shown as dark
-magenta in Figure~\ref{fig:workflow}.
+The system first identifies the dark central circular blocked region
+to act as a relative landmark for subsequent operations. Consistent
+with experimental design, we assume it contains the center of the
+diffraction pattern. The edge of the background surrounding the block
+is the brightest region, so the system first splits the image between
+areas with values less than and greater than two standard deviations
+above the mean. This partitioning yields a binary image where the
+center blocked region is surrounded by a halo of the upper end of the
+pattern background and occasional dots where diffraction peaks rise
+more than two standard deviations above background. We convert this
+binary image to a hierarchical contour set with OpenCV. We then take
+the blocked region to be the inner-most contour and model it as the
+smallest enclosing circle, shown as dark magenta in
+Figure~\ref{fig:workflow}.
 
-\dave{I cut this down and simplified. It can be removed or further
-reworked if desired.}
-With the central blocked region located, we identify local maxima in 3
-by 3 pixel groups after Gaussian smoothing and mask those found in
-regions unlikely to provide peaks of interest.  Resulting maxima are
-accepted or rejected as peaks based on masking parameters.  Masked
-rejection-regions consist of: 1) a circular zone around the central
-blocked region where the blocking generates non-peak local maxima, 2)
-areas below the 80th percentile where detector noise dominates, 3)
-areas near the image edge where peaks are partially cropped. The
-resulting unmasked area from which we keep maxima is depicted as light
-purple overlay in Figure~\ref{fig:workflow}.
+With the central blocked region located, we identify local maxima in
+three by three groups after Gaussian smoothing with kernel having a
+three pixel standard deviation. We reject resulting maxima in regions
+unlikely to provide peaks of interest before attempting to match peak
+pairs. Masked rejection-regions consist of: 1) a circular zone around
+the central blocked region where the blocking generates non-peak local
+maxima, 2) areas below the 80th percentile where detector noise
+dominates, 3) areas near the image edge where peaks are partially
+cropped. The resulting unmasked area from which we keep maxima is
+depicted as light purple overlay in Figure~\ref{fig:workflow}.
 
@@
 
-Next, maxima are clustered into peak pairs based on their distance
-and angle from the center of the blocked region. Starting with those
-maxima nearest the blocked region, a corresponding maxima is sought an
-equal distance away from the blocked region and located such that the
+Next, we cluster maxima into peak pairs based on their distance and
+angle from the center of the blocked region. Starting with those
+maxima nearest the blocked region, a corresponding maximum is sought
+an equal distance away from the blocked region and located so that the
 angle formed by the two maxima and the center of the blocked region is
-180$^\circ$. If no matched maxima exists within a 10\% tolerance, the
-maxima is discarded as a spurious peak. With peak pairs now identified
-(shown as color matched dots in Figure~\ref{fig:workflow}), the
-diffraction center is identified by taking the mean location between
-peak pairs and the background is fit and subtracted.
+180$^\circ$. If no matching maximum exists within a 10\% tolerance,
+the prior maximum is discarded as a spurious peak. With peak pairs now
+identified (shown as color matched dots in Figure~\ref{fig:workflow}),
+the diffraction center is identified by taking the mean location
+between peak pairs and the background is fit and removed.
 
-The background is subtracted by first masking arcs encompassing peak
+We subtract the background by first masking arcs encompassing peak
 pairs and then fitting a double exponential to a radial profile of the
@@
 We perform an initial peak fitting by residual minimization between a
-generated peak and the extracted ROI. This gives us a set of starting
+generated peak and the extracted ROI. This gives a set of starting
 parameters that we use, with random variation, to initialize the
@@
 challenging images are shown in Figures~\ref{fig:imgs}A and
-\ref{fig:imgs}B. Our technique segments these images into regions of
-interest, identifies the approximate locations of the diffraction
-peaks, subtracts background from the regions surrounding the peaks,
-and computes the distribution of possible parameters which underly the
-shown peaks.  We find that our initial image segmentation step
-successfully models the center blocked region of each image in greater
-than 99\% of images in our test corpus of 1,220 images. The overall
-process allows us to calculate peak-to-peak distances to sub-pixel
-accuracy with a confidence interval of 90\%. MCMC sampling combined
-with image segmentation allows us to precisely, accurately, and
-automatically locate the $d_{10}$ peak center and thus calculate the
-lattice spacing measured by a diffraction image to within 0.03 nm.
+\ref{fig:imgs}B.  Our initial image segmentation step successfully
+identifies the center blocked region of each image in greater than
+99\% of images in our test corpus. The overall process allows us to
+calculate peak-to-peak distances to sub-pixel accuracy with a
+confidence interval of 90\%. MCMC sampling combined with image
+segmentation allows us to precisely, accurately, and automatically
+locate the $d_{10}$ peak centers and thus calculate the lattice
+spacing measured by a diffraction image to within 0.03 nm.
 
@@
 
+The increasing prevalence of advanced diffractive imaging techniques
+demands data analytic methods that can handle, objectively and
+reproducible, significant increases in the volume of data generated
+and analyzed. This toolchain is a first step in a process that will
+greatly increase the power and capabilities of a widely used approach
+to understanding protein structure in living tissue.
+
 %%%%%%%%%%%

### 2015-05-29 paper.tex
Tiny edit

@@
 We apply our workflow to a test corpus of 1,220 images generated using
-X-ray diffraction during insect flight muscle research at Argonne
-National Laboratory's BioCAT Beamline. Sample high-quality and
+X-ray diffraction during insect flight muscle research at the Argonne
+National Laboratory BioCAT Beamline. Sample high-quality and
 challenging images are shown in Figures~\ref{fig:imgs}A and

### 2015-06-03 paper.tex
Swapped paper format to LNCS

@@
-\documentclass{sig-alternate}
+\documentclass{llncs}
 
 \usepackage[usenames,dvipsnames]{color}
-\usepackage{xspace}
+\usepackage{graphicx}
 
-\usepackage{fancyvrb}
-\DefineVerbatimEnvironment{code}{Verbatim}{fontsize=\small}
-\DefineVerbatimEnvironment{example}{Verbatim}{fontsize=\small}
-
-% Looks better (and more concise) than Times New Roman
-\usepackage{times}
-% Better spacing
-\usepackage{microtype}
-\usepackage[normalem]{ulem}
-\usepackage{enumitem}
-% Caption package both lets you set the spacing between figure and caption
-% and also makes the \figref{} point to the right place.
-\usepackage[font=bf,aboveskip=0pt,belowskip=-12pt]{caption}
-% Get the caption package to work with ACM style
-\DeclareCaptionType[placement=b,within=none]{copyrightbox}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\newfont{\mycrnotice}{ptmr8t at 7pt}
-\newfont{\myconfname}{ptmri8t at 7pt}
-\let\crnotice\mycrnotice%
-\let\confname\myconfname%
-
-\permission{Permission to make digital or hard copies of all or part of this work for personal or classroom use is granted without fee provided that copies are not made or distributed for profit or commercial advantage and that copies bear this notice and the full citation on the first page.
-Copyrights for components of this work owned by others than the author(s) must be honored.
-Abstracting with credit is permitted.
-To copy otherwise, or republish, to post on servers or to redistribute to lists, requires prior specific permission and/or a fee.
-Request permissions from Permissions@acm.org.}
-\conferenceinfo{}{DMAH 2015, September 4th, Kohala Coast, USA}
-\copyrightetc{}
-\crdata{}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 % Allow comments by name
-\newcommand{\note}[2]{
-    \textbf{\textcolor{#1}{#2}}
-}
+\renewcommand{\note}[2]{\textbf{\textcolor{#1}{#2}}}
 \newcommand{\comment}[1]{\note{red}{#1}}
@@
 \newcommand{\tom}[1]{\note{OrangeRed}{Tom: #1}}
-
 % To remove all notes, uncomment the next line.
 %\renewcommand{\note}[2]{\unskip}
-
-\newcommand{\ie}{{\em i.e.}\ }
-\newcommand{\eg}{{\em e.g.}\ }
-\newcommand{\ea}{{\em et al.}\xspace}
-\newcommand{\aka}{{\em a.k.a.}\ }
-
-\usepackage{url}
-\usepackage[pdftex]{hyperref}
 
@@
   \includegraphics[width=\linewidth]{figures/x_ray_image_montage}
-  \vspace{-12pt}
   \caption{\label{fig:imgs}
@@
 	}
-	\vspace{-2pt}
+	\vspace{-10pt}
 \end{figure}
@@
   \includegraphics[width=\linewidth]{figures/img_analysis}
-  \vspace{-8pt}
   \caption{\label{fig:workflow}
@@
 \begin{document}
+
+%%%%%%%%%%%
+\frontmatter
+\pagestyle{headings}  % switches on printing of running heads
+\addtocmark{MCMC Analysis of X-ray Diffraction Images} % additional mark in the TOC
+
+%%%%%%%%%%%
+\mainmatter
 \title{Automated Analysis of Muscle X-ray Diffraction \\ Imaging with MCMC}
-\author{C.\ David Williams$^{1}$,  Magdalena Balazinska$^2$, Thomas Daniel$^1$ \\
-\affaddr{$^1$Department of Biology, $^2$Department of Computer Science \& Engineering}\\
-\affaddr{University of Washington} \\
+\author{C.\ David Williams\inst{1} \and Magdalena Balazinska\inst{1} \and Thomas Daniel\inst{1}}
+\institute{Department of Biology, University of Washington, Seattle WA 98195, USA, \\
+  \email{cdave@uw.edu}
   }
@@
 
-% \dave{Presumably this gets filled in later?}
-%    \vspace{-8pt}
-%    \category{H.2.4}{Database Management}{Systems---Distributed databases, Query processing, Relational databases}
-%    \vspace{-8pt}
-%    \keywords{Myria; parallel data management; Cloud service; astronomy}
+%%%%%%%%%%%
+\begin{abstract}
 
-\sloppy
-
-%%%%%%%%%%%
-%\section{Questions addressed}
-%\label{sec:addressed}
-
-%In this work, we focus on the following questions:
-
-%\begin{itemize}[noitemsep]
-%\item Are sufficient conserved markers available in small-angle scattering diffraction images for automatic segmentation?
-%\item Does MCMC distribution fitting provide reliable estimates of diffraction peak properties such as center and decay?
-%\end{itemize}
+  High-speed X-ray diffraction is the state-of-the-art approach to
+  understanding protein structure and dynamics in living tissues,
+  especially muscles.  Existing analytic approaches, however, require
+  expert hand-digitization to extract parameters of interest. This
+  produces repeatable measurements, but remains subjective and does
+  not offer information on the precision of the measured parameters or
+  strict reproducibility of analyzed data.  We developed a processing
+  tool chain, which first segments the diffraction image into regions
+  of interest using highly conserved features and then samples the
+  possible parameter values with a Markov chain Monte Carlo approach.
+  Our approach produces an automated, reproducible, objective estimate
+  of relevant image parameters.
+  \keywords{image analysis, X-ray diffraction, muscle structure, MCMC modeling}
+\end{abstract}
 
@@
 
-\begin{itemize}[noitemsep]
+\begin{itemize}
 \item Development of a declarative language to describe processing

### 2015-07-23 paper.tex
Camera ready

@@
 % Figures
+\newcommand{\figurerig}{
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=.8\linewidth]{figures/rig_and_lattice}
+  \caption{\label{fig:rig}
+    Data collection and interpretation.  
+    A) Diffractive imaging of an isolated muscle sample requires a
+    physiological solution bath with top entry for mounting to
+    force/length controllers (arms) and front and rear cut outs to
+    allow the X-ray beam to pass through and scatter off the muscle
+    sample. B) Muscle's contractile lattice is composed of thick and
+    thin filaments. When viewed down the long axis, these filaments
+    form a hexagonal array. The $d_{10}$ spacing is the distance
+    between adjacent lines of thick filaments, and is proportional to
+    the distance molecular motors (gray links) must diffuse across in
+    order to bind. 
+	}
+	\vspace{-10pt}
+\end{figure}
+}
+
 \newcommand{\figureimgs}{
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/x_ray_image_montage}
+  \includegraphics[width=.8\linewidth]{figures/x_ray_image_montage}
   \caption{\label{fig:imgs}
@@
     multiple $d_{10}$ peaks which must be segregated during the
-    detection process.  
+    detection process. Such two-fiber exposures occur rarely in the 
+    test corpus used in this work and are successfully processed by 
+    the automated grouping of peaks. 
 	}
@@
   	Analysis workflow.
-    The blocked center region is detected and used as a landmark.  Low
-    signal regions and edges are masked. The image is smoothed and
-    peaks are detected, clustered, and classified. The background
-    (sans that surrounding diffraction lines) is computed and
-    subtracted. Regions of interest surrounding the $d_{10}$ peaks are
-    extracted and the generating distributions are fit with an MCMC
-    sampler. We marginalize across peak parameters other than those of
-    interest.  
+    A) An X-ray diffraction image is read in. B) The blocked center
+    region (in green) is detected and used as a landmark.  Low signal
+    regions, the central blocked region, and image edges are masked
+    off, leaving only the foreground outlined in white. C) The image
+    is smoothed and peaks are detected, paired, and classified. To
+    remove the exponentially decaying background from the peaks we
+    mask off the area around the diffraction lines, shown in gray. The
+    remaining background is then collapsed into a radial profile
+    centered at the mean peak position.  A double exponential fit to
+    this profile is computed and subtracted from the regions of
+    interest surrounding the $d_{10}$ peaks, leaving just the peak
+    signal shown in D).  E) We use an MCMC sampler to calculate the
+    parameters generating the distributions these peaks are sampled
+    from. F) We marginalize across peak parameters other than those of
+    interest, extracting peak positions and the error in those
+    position estimations.  
+	}
+	\vspace{-10pt}
+\end{figure}
+}
+
+\newcommand{\figureerror}{
+\begin{figure}[tbp]
+  \centering
+  \includegraphics[width=\linewidth]{figures/hand_vs_auto}
+  \caption{\label{fig:error}
+  	Validation against hand-measurement. 
+    A) The majority of residuals (between the manually and
+    automatically measured $d_{10}$ peak distances) are within a
+    single pixel of the manually digitized validation values. In a
+    handful of outliers the automated method mis-identifies a closer
+    structure as the $d_{10}$ pair. B) The cumulative distribution of
+    the residuals shows that half of the automated measurements agree
+    within 0.8 pixels and over 90\% agree within 3 pixels. A quartile
+    box plot above shows the heavily left-weighted asymmetry of the
+    residual distribution.  
     }
@@
 \title{Automated Analysis of Muscle X-ray Diffraction \\ Imaging with MCMC}
-\author{C.\ David Williams\inst{1} \and Magdalena Balazinska\inst{1} \and Thomas Daniel\inst{1}}
-\institute{Department of Biology, University of Washington, Seattle WA 98195, USA, \\
+\author{C.\ David Williams\inst{1} \and Magdalena Balazinska\inst{1} \and Thomas L.\ Daniel\inst{1}}
+\institute{University of Washington, Seattle WA 98195, USA, \\
   \email{cdave@uw.edu}
@@
 lattice of interacting elastic and force generating molecules to
-create controlled macro-scale movement \cite{Millman1998}. The advent
-of advanced X-ray diffraction imaging lets us analyze protein motions
-at spatial and temporal scales never previously realized, giving novel
-insight into the molecular basis of motion in living systems.
+create controlled macro-scale movement \cite{Reedy1965,Millman1998}.
+The advent of advanced X-ray diffraction imaging lets us analyze
+protein motions at spatial and temporal scales never previously
+realized, giving novel insight into the molecular basis of motion in
+living systems.
+
+Exploring the dynamics of protein interactions in muscle requires
+imaging that provides information at the nanometer to Angstrom scale.
+X-ray diffraction provides that information and is unique in its
+ability to image live muscle during movement, where other techniques
+(e.g.\ solution biochemistry or cryo electron microscopy) are only able
+to resolve properties of isolated individual molecular motors or
+anatomic features of static, dried muscle. This ability to image
+\textit{in vivo} makes X-ray diffraction the only direct means of
+measuring the structural changes that generate force; all other
+techniques that permit some level of molecular motor activity require
+the removal of those motors from the regulating structure of the
+thick/thin filament lattice and are thus less informative about the
+processes that actually control activation.  This is a key advantage
+of X-ray diffraction: changes in muscle microstructure are highly
+correlated with changes in force production and any attempt to
+separate the two removes significant aspects of the system's response
+to real-world stimuli \cite{BrennerYu1985,Bagni1994}.
 
 Surprisingly, we still use human experts to manually extract
-structural parameters from X-ray images with NIH ImageJ. This produces
+structural parameters from X-ray images with NIH ImageJ via manual
+selection of peaks and rock-of-eye parameter fitting. This produces
 repeatable measurements but is subjective, fails to provide confidence
@@
 analytic technique, with a single digitizer able to process only a few
-hundred images a day.  Historically this rate has been sufficient, but
-new high-speed imaging systems let us investigate short-timescale
+hundreds of images a day.  Historically this rate has been sufficient,
+but new high-speed imaging systems let us investigate short-timescale
 components of muscle contraction and generate data sets with many
@@
 
+\figurerig
+
 We seek to build a service for the automated analysis of muscle
@@
 using a declarative query interface and the system should
-automatically process the user's image database. In this extended
-abstract, we present the first components of the processing tool chain
-at the heart of this service. The toolchain first segments the
+automatically process the user's image database. In this work, we
+present the first components of the processing tool chain at the heart
+of this service. The toolchain first segments the
 diffraction image into regions of interest using conserved features
@@
 
-We initially focus on the $d_{10}$ parameter that determines the
-distance which muscle's molecular motors must bridge in order to bind
-and generate force \cite{Williams2013}. This distance changes during
-contraction and thus regulates the force produced.
+We focus initially on measuring the $d_{10}$ parameter, a crucial
+spacing in muscle shown in Figures~\ref{fig:rig} and \ref{fig:imgs}.
+The $d_{10}$ spacing determines the distance which muscle's molecular
+motors must bridge in order to bind and generate force
+\cite{Williams2013}. This distance changes during contraction and thus
+regulates the force produced \cite{George2013}.
 
@@
 \subsection{Image segmentation}
+\label{subsec:imgseg}
 
@@
 more than two standard deviations above background. We convert this
-binary image to a hierarchical contour set with OpenCV. We then take
-the blocked region to be the inner-most contour and model it as the
-smallest enclosing circle, shown as dark magenta in
-Figure~\ref{fig:workflow}.
+binary image to a hierarchical contour set with OpenCV, an open source
+library of standard image processing techniques \cite{opencv}. We then
+take the blocked region to be the inner-most contour and model it as
+the smallest enclosing circle, shown as light green in
+Figure~\ref{fig:workflow}B.
 
@@
 cropped. The resulting unmasked area from which we keep maxima is
-depicted as light purple overlay in Figure~\ref{fig:workflow}.
+bounded by a white border in Figure~\ref{fig:workflow}B.
 
@@
 angle formed by the two maxima and the center of the blocked region is
-180$^\circ$. If no matching maximum exists within a 10\% tolerance,
-the prior maximum is discarded as a spurious peak. With peak pairs now
-identified (shown as color matched dots in Figure~\ref{fig:workflow}),
-the diffraction center is identified by taking the mean location
-between peak pairs and the background is fit and removed.
+180$^\circ$. In cases where the matching maxima is clipped by the
+image frame or fails to be distinguished from background the initial
+maximum is discarded. For a maxima to be considered a match its
+distance and angle must match those of the initial maxima within 10\%.
+With peak pairs now identified (shown as color matched dots in
+Figure~\ref{fig:workflow}C), the diffraction center is identified by
+taking the mean location between peak pairs and the background is
+removed.
 
@@
 (shown as a light gray arc under the peak pairs in
-Figure~\ref{fig:workflow}). We calculate a radial profile of the image
+Figure~\ref{fig:workflow}C). We calculate a radial profile of the image
 around the pattern center, omitting diffraction line regions. We fit a
@@
 an estimated background image and subtract it from the real image,
-allowing us to extract the $d{10}$ peaks as regions of interest (ROIs)
+allowing us to extract the $d_{10}$ peaks as regions of interest (ROIs)
 unhindered by an overlaid diffraction background.
@@
 \subsection{Image modeling with MCMC processes}
+\label{subsec:modelmcmc}
 
@@
 being drawn from an underlying Pearson VII distribution, commonly used
-to fit X-ray diffraction peaks. This process allows us to generate
-possible peak matches using five parameters: peak center x-location,
-peak center y-location, peak height, peak spread, and peak decay rate.
-We perform an initial peak fitting by residual minimization between a
-generated peak and the extracted ROI. This gives a set of starting
-parameters that we use, with random variation, to initialize the
-positions of the MCMC agents that will explore the model space.
+to fit X-ray diffraction peaks \cite{Hall1977}. This process allows us
+to generate possible peak matches using five parameters: peak center
+x-location, peak center y-location, peak height, peak spread, and peak
+decay rate.  We perform an initial peak fitting by residual
+minimization between a generated peak and the extracted ROI. This
+gives a set of starting parameters that we use, with random variation,
+to initialize the positions of the MCMC agents that will explore the
+model space.
 
@@
 center to precisely calculate $d_{10}$ spacing (as in
-Figure~\ref{fig:workflow}), we automatically integrate our uncertainty
+Figure~\ref{fig:workflow}F), we automatically integrate our uncertainty
 about peak height, spread, and decay.
@@
 challenging images are shown in Figures~\ref{fig:imgs}A and
-\ref{fig:imgs}B.  Our initial image segmentation step successfully
-identifies the center blocked region of each image in greater than
-99\% of images in our test corpus. The overall process allows us to
-calculate peak-to-peak distances to sub-pixel accuracy with a
-confidence interval of 90\%. MCMC sampling combined with image
-segmentation allows us to precisely, accurately, and automatically
-locate the $d_{10}$ peak centers and thus calculate the lattice
-spacing measured by a diffraction image to within 0.03 nm.
+\ref{fig:imgs}B \cite{Irving2006}. We tested the output of our
+automated system against the results from a domain expert manually
+digitizing peak locations. Our initial image segmentation step,
+described in Section~\ref{subsec:imgseg}, successfully identifies the
+center blocked region of each image in greater than 99\% of images in
+our test corpus. The overall process allows us to calculate
+peak-to-peak distances to sub-pixel accuracy with a confidence
+interval of 90\%, with residuals between expectation values and
+manually digitized peak positions shown in Figure~\ref{fig:error}.
+MCMC sampling combined with image segmentation allows us to precisely,
+accurately, and automatically locate the $d_{10}$ peak centers and
+thus calculate the lattice spacing measured by a diffraction image to
+within 0.03 nm.
+
+\figureerror
 
@@
     of the analysis pipeline will facilitate adoption. 
+\item This toolchain works best when applied to sharp and well defined
+    images produced by isolated fibers. When working with images
+    produced by \textit{in vivo} samples, such as those taken through
+    a fly thorax, diffraction from multiple muscles in different
+    orientations is common. These multiple-muscle images produce a
+    more extreme version of the image shown in Figure~\ref{fig:imgs}B,
+    where twin rotated peaks that are twisted and blurred about the
+    center of the diffraction pattern and may have substantially
+    divergent $d_{10}$ distances. Twin rotated peaks must be classified by
+    diffraction line/muscle and treated independently during analysis. 
 \item Application of these techniques to coming ultra-high
@@
 %%%%%%%%%%%
-%\section{Acknowledgments}
-%\label{sec:acknowledgments}
-%
-%This work was supported in part by the Army Research Office through
-%ARO Grants W911NF-13-1-0435 and W911NF-14-1-0396, an award from 
-%the Gordon and Betty Moore Foundation and the Alfred P Sloan 
-%Foundation, the Washington Research Foundation Fund for Innovation 
-%in Data-Intensive Discovery, and the UW eScience Institute.
-%
-%We thank Jake VanderPlas for helpful discussions of statistical
-%techniques, Tom Irving for advice on X-ray imaging, and Simon Sponberg
-%for the sharing of diffraction images.
+\section{Acknowledgments}
+\label{sec:acknowledgments}
+
+This work was supported in part by NSF grant IIS-1110370, the Intel
+Science and Technology Center for Big Data, the Army Research Office
+through ARO Grants W911NF-13-1-0435 and W911NF-14-1-0396, an award
+from the Gordon and Betty Moore Foundation and the Alfred P Sloan
+Foundation, the Washington Research Foundation Fund for Innovation in
+Data-Intensive Discovery, and the UW eScience Institute.
+
+We thank Jake VanderPlas for helpful discussions of statistical
+techniques, Tom Irving for advice on X-ray imaging, Gideon Dunster for
+digitization of diffraction images, and Simon Sponberg for the sharing
+of diffraction images.
 
